Without loss of generality, \\ any reduced ring is a field

### 2020-04-02 paper-herrsching2019/main.tex
Generic models and more

\documentclass{ws-rv9x6}
\usepackage[onethmnum]{ws-rv-thm}
\usepackage{ws-rv-van}
\usepackage{mathtools,stmaryrd,xspace}

@@

There is nothing special about the real numbers in this example;
in fact, the example works just as well with any set~$M$ in place of the reals.
For any set~$M$, there is a locale of surjections~$\NN \twoheadrightarrow M$,
and this locale is trivial (isomorphic to the empty locale) if and only if~$M$
is empty. Locales of these kind are used as an important reduction step in the
extension of Grothendieck's Galois theory by Joyal and
Tierney~\cite[Section~V.3]{joyal-tierney:galois-theory}.

@@

\begin{exercise}{Local locales}%
A locale~$X$ is \emph{local} iff for any open covering~$\top = \bigvee_i U_i$,
there is an index~$i$ such that~$\top = U_i$. (For instance, the spectrum of a
ring~$A$ is local iff~$A$ is local.)
Show that any local locale has a canonical point, the \emph{focal point}.
\end{exercise}

@@

\begin{definition}A \emph{geometric theory} consists of
\begin{enumerate}
  \item a set of sorts: $X$, $Y$, $Z$, \ldots
  \item a set of function symbols: $f : X \times Y \to Z$, \ldots
  \item a set of relation symbols: $R \hookrightarrow X \times Y \times Z$, \ldots
  \item a set of geometric sequents as axioms: $\varphi \vdash_{x:X, y:Y} \psi$, \ldots
\end{enumerate}
A \emph{geometric sequent} (in some context~$x_1\?X_1,\ldots,x_n\?X_n$) is a
formula built using only the ingredients ${=}\ {\top}\ {\wedge}\ {\bot}\
{\vee}\ {\bigvee}\ {\exists}$ and the relation symbols (but no ${\Rightarrow}\
{\forall}$). The symbol~``$\bigvee$'' refers to disjunctions of arbitrary
set-indexed families of formulas.
\end{definition}

@@

@@

The question whether a given locale is spatial is intimately related with the
existence of appropriate \emph{ideal objects} whose existence typically hinges
on the axiom of choice or one of its variants. For instance, there are the
following results (where we allow ourselves a modicum of classical logic in
order to state the results in a familiar form).
\begin{enumerate}
\item The localic spectrum of a ring~$A$ is spatial iff every nontrivial ideal
(of a certain class of rings related with~$A$) is contained in a prime ideal.
\item The localic Galois group of a Galois extension~$L|k$ is spatial iff
any~$k$-homomorphism~$\sigma : E \to L$ defined on a finite intermediate
extension~$L|E|k$ admits an extension to a~$k$-homomorphism~$L \to L$.
\item The localic real line is spatial iff every open covering of the
metric space~$[0,1]$ of Dedekind reals has a Lebesgue number (that is, if the
metric space~$[0,1]$ is Heine--Borel compact).
\item The statement that the classifying locale of any propositional coherent
theory (that is, a propositional geometric theory in which only finite
disjunctions occur) is spatial is equivalent to the Boolean Prime Ideal
Theorem~\BPIT.
\end{enumerate}
These results explain why in classical mathematics, where all of these
conditions are satisfied, we can blithely and fruitfully use the topological
spectrum of a ring, the topological Galois group or the topological space of
reals. In classical mathematics, employing the localic counterparts is merely
for convenience or for aesthetic reasons; in constructive mathematics, the
localic replacements (or similar replacements in other pointfree approaches to
topology) are vital.

%\begin{theorem}The localic real line is spatial if and only if the set~$[0,1]$ of
%Dedekind real numbers is compact in the sense of Heine--Borel.
%\end{theorem}
%
%\begin{proof}XXX\end{proof}

%\begin{theorem}Let~$A$ be a ring and~$\Spec(A)$ its spectrum as defined above as the locale of
%prime filters of~$A$. Then~$\Spec(A)$ is spatial if (the law of excluded middle and) the
%Boolean Prime Ideal Theorem~\BPIT hold. More specifically, $\Spec(A)$ is
%spatial if and only if for any ideal~$\aaa \subseteq A$ and any element~$x \in A$,
%\[ x \in \sqrt{\aaa} \quad\textnormal{if}\quad
%\textnormal{for all filters~$\fff \subseteq A$ with~$x \in \fff$, $\fff \between \aaa$}. \]
%\end{theorem}
%
%The symbol~``$\between$'' denotes that the two sets have an element in common.
%
%\begin{proof}Assume that~$\Spec(A)$ is spatial. Let an ideal~$\aaa \subseteq A$
%and an element~$x \in A$ be given such that for all filters~$\fff \subseteq A$
%with~$x \in \fff$, $\fff \between \aaa$. Then for all filters~$\fff$, $\fff
%\inplus \alpha_x \Rightarrow \fff \inplus \bigvee_{y\in \aaa} \alpha_y$.
%Hence ...
%Using the filter axioms, any open of~$\Spec(A)$ can be written
%as~$[\bigvee_i \alpha_{x_i}]$.
%\end{proof}

@@

Definition~\ref{defn:presheaf}, as well as any other definition in this section,
also makes sense for topological spaces instead of locales, since only the
notion of opens and their inclusion relation is used. The elements of~$F(U)$ are
called \emph{sections of~$F$ over~$U$} and the maps~$(\cdot)|^U_V$
are called \emph{restriction maps}.

@@

Let~$X$ be a locale. A \emph{formula over an open~$U \in \O(X)$} is a
first-order formula (made up using~${=}\ {\top}\ {\wedge}\ {\bot}\ {\vee}\
{\Rightarrow}\ {\forall}\ {\exists}$) over the signature which has one sort for
each sheaf~$F$, one constant symbol of sort~$F$ for each section~$s \in F(U)$,
one function symbol~$f : F \to G$ for each morphism of sheaves, and so on.

Table~\ref{table:sheaf-semantics} defines by recursion on the structure of
formulas what it means for a formula~$\varphi$ over an open~$U$ to be
\emph{forced} or \emph{true over~$U$}, written~``$U \models \varphi$''. The
main differences with the perhaps more familiar Kripke semantics are marked.

@@

\begin{definition}A \emph{sheaf model}~$M$ of a geometric theory~$\TT$ over a
locale~$X$ consists of
\begin{enumerate}
  \item a sheaf~$\brak{S}$ on~$X$ for each sort~$S$,
  \item a morphism~$\brak{f} : \brak{S_1} \times \cdots \times \brak{S_n} \to
  \brak{T}$ of sheaves for each function symbol~$f : S_1 \times \cdots \times S_n \to T$ and
  \item a subsheaf~$\brak{R} \subseteq \brak{S_1} \times \cdots \times \brak{S_n}$
  for each relation symbol~$R \hookrightarrow S_1 \times \cdots \times S_n$
\end{enumerate}
such that~$M$ validates the axioms of~$\TT$, that is, such that the axioms
of~$\TT$ are forced on~$\top \in \O(X)$.
\end{definition}

@@

\begin{exercise}{The localic Galois group of a field extension}%
Let~$L|k$ be a field extension. The \emph{localic Galois group}~$\Gal(L|k)$
of~$L|k$ is the classifying locale of the theory which has one nullary relation
symbol~``$\sigma_{xy}$'' for each pair~$\langle x,y \rangle \in L$ and the
following axioms:
\begin{align*}
  \top &\vdash \bigvee_y \sigma_{xy} & \text{(for each~$x \in L$)} \\
  \top &\vdash \bigvee_x \sigma_{xy} & \text{(for each~$y \in L$)} \\
  \sigma_{xy} \wedge \sigma_{xy'} &\vdash \bigvee\{ \top \,|\, y = y' \} & \text{(for each~$x,y,y' \in L$)} \\
  \sigma_{xy} \wedge \sigma_{x'y} &\vdash \bigvee\{ \top \,|\, x = x' \} & \text{(for each~$x,x',y' \in L$)} \\
  \top &\vdash \sigma_{00} \\
  \top &\vdash \sigma_{11} \\
  \sigma_{xy} \wedge \sigma_{x'y'} &\vdash \sigma_{x+x',y+y'} & \text{(for each~$x,x',y,y' \in L$)} \\
  \sigma_{xy} \wedge \sigma_{x'y'} &\vdash \sigma_{xx',yy'} & \text{(for each~$x,x',y,y' \in L$)} \\
  \top &\vdash \sigma_{xx} & \text{(for each~$x \in k$)}
\end{align*}
\begin{enumerate}
\item Explain how to obtain a morphism~$\Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send an automorphism~$\sigma$ to its inverse~$\sigma^{-1}$.
\item Explain how to obtain a morphism~$\Gal(L|k) \times \Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send a pair~$\langle \sigma, \tau \rangle$ to~$\sigma \circ \tau$.
\item Assume from now on that~$L$ is a \emph{geometric field} in the sense that any
element is zero or invertible. Show that the axiom expressing injectivity can
be dropped without changing the set of provable sequents.
\item Let~$f \in k[X]$ be a monic polynomial such that~$f(X) =
(X-x_1)\cdots(X-x_n)$ over~$L$. Let~$x$ be one of the zeros of~$f$. Show that
the theory shows~$\top \vdash \bigvee_{i=1}^n \sigma_{x,x_i}$.
\item Now furthermore assume that~$L$ is \emph{Galois over}~$k$ in the sense that
any element of~$L$ is the zero of a monic polynomial~$f \in k[X]$ which is
separable (that is,~$(f,f') = (1) \subseteq k[X]$) and splits over~$L$ into
linear factors. Show that the axioms expressing surjectivity can be dropped.
\item Show that the topological space of points of~$\Gal(L|k)$ is canonically
homeomorphic to the topological Galois group of~$L|k$.
\item Show that~$\Gal(L|k)$ is spatial in classical mathematics.
\end{enumerate}
\end{exercise}

@@

\begin{enumerate}
\item Instead of replacing a given ring with another (for instance a ring~$A$ with
one of its stalks~$A_\ppp$ at a minimal prime ideal), we replace a given ring
with a \emph{sheaf of rings}.
\item We then maintain the illusion that we are working with a plain old ring
by employing the sheaf semantics.
\end{enumerate}

@@

The locale~$X$ on which this theater plays out is none other than the
\emph{spectrum of~$A$} from algebraic geometry, and the sheaf~$A^\sim$ is the
sheaf of regular functions. We will give details on its construction in
Section~\ref{sect:origins}, but first want to reflect on the displayed
statements.

@@

\paragraph{The field property.}
The fact that~$A^\sim$ is a field is where~$A^\sim$ derives its main usefulness
from. While statements~(\ref{item:loc-stable}), (\ref{item:stalks}) and
(\ref{item:local}) are immediate consequences of the way~$A^\sim$ is
constructed (Proposition~\ref{prop:immediate-consequences}), the field property~(\ref{item:field})
requires an actual computation for its verification and was a surprising
discovery by Mulvey in the 1970s.

Tierney commented around that time that the field property ``is surely important,
though its precise significance is still somewhat obscure---as is the case with
many such nongeometric formulas''~\cite[p.~209]{tierney:spectrum}. Even though
the property has been known for a long time, only recently was it put in a
wider context (recognizing it as a small shadow of algebraic geometry's
\emph{quasicoherence}) and its usefulness for constructive algebra
appreciated~\cite[Sections~3.3 and~11.4]{blechschmidt:phd}.

@@

In exactly the same manner, we can construct the free group over a monoid, the
free~$\ZZ/(2)$-algebra over a ring, the free ring over a set or the free
Heyting algebra over a partially ordered set. This is all well-understood; a
particularly perspicious account in the general setting of partial Horn
theories is due to Palmgren and Vickers~\cite{palmgren-vickers:partial-horn}.

@@

\begin{exercise}{Bridging modules and their induced sheaves}%
Let~$M$ be a module over a ring~$A$.
\begin{enumerate}
\item Show that~$M$ is finitely generated iff~$M^\sim$ is.
\item Show that~$M$ is finite locally free (that is, that there is a
partition~$1 = f_1 + \cdots + f_n$ such that each module~$M[f_i^{-1}]$ is
finite free over~$A[f_i^{-1}]$) iff~$M^\sim$ is finite free.
\end{enumerate}
\end{exercise}

@@

@@

>>>>

\documentclass{ws-rv9x6}
\usepackage[onethmnum]{ws-rv-thm}
\usepackage{ws-rv-van}
\usepackage{mathtools,stmaryrd,xspace,bussproofs}

@@

There is nothing special about the real numbers in this example;
in fact, the example works just as well with any set~$M$ in place of the reals.
For any set~$M$, there is a locale of surjections~$\NN \twoheadrightarrow M$,
and this locale is trivial (isomorphic to the empty locale) if and only if~$M$
is empty. Locales of these kind are used as an important reduction step in the
extension of Grothendieck's Galois theory by André Joyal and
Myles Tierney~\cite[Section~V.3]{joyal-tierney:galois-theory}.

@@

\begin{exercise}{Local locales}%
A locale~$X$ is \emph{local} iff for any open covering~$\top = \bigvee_i U_i$,
there is an index~$i$ such that~$\top = U_i$. (For instance, the spectrum of a
ring~$A$ is local iff~$A$ is local.)
Show that any local locale has a canonical point, its \emph{focal point}.
\end{exercise}

@@

\begin{definition}A \emph{geometric theory} consists of
\begin{enumerate}
  \item a set of sorts: $X$, $Y$, $Z$, \ldots
  \item a set of function symbols: $f : X \times Y \to Z$, \ldots
  \item a set of relation symbols: $R \hookrightarrow X \times Y \times Z$, \ldots
  \item a set of geometric sequents as axioms: $\varphi \vdash_{x:X, y:Y} \psi$, \ldots
\end{enumerate}
A \emph{geometric sequent} (in some context~$x_1\?X_1,\ldots,x_n\?X_n$) is a
formula built using only the ingredients ${=}\ {\top}\ {\wedge}\ {\bot}\
{\vee}\ {\bigvee}\ {\exists}$ and the relation symbols (but no ${\Rightarrow}\
{\forall}$). The symbol~``$\bigvee$'' refers to disjunction of arbitrary
set-indexed families of formulas. The theory \emph{proves} a sequent~$\sigma$ if and
only if there exists a \emph{derivation} of~$\sigma$, where the set of
derivations is inductively generated by the axioms and the rules displayed in
Table~\ref{table:geometric-logic}.
\end{definition}

{\makeatletter
\newcommand*{\shifttext}[1]{%
  \settowidth{\@tempdima}{#1}%
  \makebox[\@tempdima]{\hspace*{\@tempdima}#1}%
}
\makeatother
\begin{table}[t]
  \tbl{The rules of geometric logic}{\begin{minipage}{0.8\textwidth}
    \centering
    \medskip

    \textnormal{structural rules} \\\smallskip
    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \seq{\vec x} \varphi$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \psi$}\AxiomC{$\psi \seq{\vec x}
    \chi$}\BinaryInfC{$\varphi \seq{\vec x} \chi$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \psi$}\UnaryInfC{$\varphi[\vec s/\vec x]
    \seq{\vec y} \psi[\vec s/\vec x]$}\DisplayProof
    \makebox[0pt]{\shifttext{\tiny ($\vec y$ not bound)}}  % in~$\varphi$ and~$\psi$)}}
    \phantom{a}\hfill
    \bigskip

    \textnormal{rules for conjunction} \\\smallskip
    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \seq{\vec x} \top$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \wedge \psi \seq{\vec x} \varphi$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \wedge \psi \seq{\vec x} \psi$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \psi$}\AxiomC{$\varphi \seq{\vec x} \chi$}\BinaryInfC{$\varphi \seq{\vec x} \psi \wedge \chi$}\DisplayProof
    \phantom{a}\hfill
    \bigskip

    \textnormal{rules for finitary disjunction} \\\smallskip
    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\bot \seq{\vec x} \varphi$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi \seq{\vec x} \varphi \vee \psi$}\DisplayProof\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\psi \seq{\vec x} \varphi \vee \psi$}\DisplayProof\hfill
    \AxiomC{$\varphi \seq{\vec x} \chi$}\AxiomC{$\psi \seq{\vec x} \chi$}\BinaryInfC{$\varphi \vee \psi \seq{\vec x} \chi$}\DisplayProof
    \phantom{a}\hfill
    \bigskip

    \textnormal{rules for infinitary disjunction} \\\smallskip
    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}\UnaryInfC{$\varphi_{i_0} \seq{\vec x} \bigvee_{i \in I} \varphi_i$}\DisplayProof\hfill
    \AxiomC{$\varphi_i \seq{\vec x} \chi \text{ for each~$i \in I$}$}\UnaryInfC{$\bigvee_{i \in I} \varphi_i \seq{\vec x} \chi$}\DisplayProof
    \phantom{a}\hfill
    \bigskip

    \textnormal{double rule for existential quantification} \\\smallskip
    \phantom{a}\hfill
    \Axiom$\varphi \ \fCenter\seq{\vec x, y} \psi$
    \doubleLine
    \UnaryInf$\exists y\?Y\_\! \varphi\ \fCenter\seq{\vec x} \psi$
    \DisplayProof
    \makebox[0pt]{\shifttext{\tiny ($y$ not free in~$\psi$)}}
    \hfill\phantom{a}
    \bigskip

    \textnormal{rules for equality} \\\smallskip
    \vspace{-0.5em}
    \phantom{a}\hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}
    \UnaryInfC{$\top \seq{x} x = x$}
    \DisplayProof
    \hfill
    \AxiomC{$\phantom{\seq{\vec x}}$}
    \UnaryInfC{$(\vec x = \vec y) \wedge \varphi \seq{\vec z} \varphi[\vec y/\vec x]$}
    \DisplayProof
    \hfill\phantom{a} \\[0.5em]
    (``$\vec x = \vec y\,$'' is an abbreviation for~``$x_1 = y_1 \wedge \cdots \wedge x_n =
    y_n$''.)%
  \end{minipage}}
  \label{table:geometric-logic}
\end{table}}

@@

\paragraph{The geometric theory of automorphisms of a field.}\label{par:theory-automorphisms}
Let~$L|k$ be a field
extension. The classifying locale of automorphisms~$L \to L$ which fix~$k$ has
one nullary relation symbol~``$\sigma_{xy}$'' for each pair~$\langle x,y
\rangle \in L$ and the following axioms:
\begin{align*}
  \top &\vdash \bigvee_y \sigma_{xy} & \text{(for each~$x \in L$)} \\
  \top &\vdash \bigvee_x \sigma_{xy} & \text{(for each~$y \in L$)} \\
  \sigma_{xy} \wedge \sigma_{xy'} &\vdash \bigvee\{ \top \,|\, y = y' \} & \text{(for each~$x,y,y' \in L$)} \\
  \sigma_{xy} \wedge \sigma_{x'y} &\vdash \bigvee\{ \top \,|\, x = x' \} & \text{(for each~$x,x',y \in L$)} \\
  \top &\vdash \sigma_{00} \\
  \top &\vdash \sigma_{11} \\
  \sigma_{xy} \wedge \sigma_{x'y'} &\vdash \sigma_{x+x',y+y'} & \text{(for each~$x,x',y,y' \in L$)} \\
  \sigma_{xy} \wedge \sigma_{x'y'} &\vdash \sigma_{xx',yy'} & \text{(for each~$x,x',y,y' \in L$)} \\
  \top &\vdash \sigma_{xx} & \text{(for each~$x \in k$)}
\end{align*}
A model of this theory (can be reorganized to be) an element of the classical
Galois group of~$L|k$.

@@

The question whether a given locale is spatial is intimately related with the
existence of appropriate \emph{ideal objects} whose existence typically hinges
on the axiom of choice or one of its variants. For instance, there are the
following results (where we allow ourselves a modicum of classical logic in
order to state the results in a familiar form).
\begin{enumerate}
\item The localic spectrum of a ring~$A$ is spatial iff every nontrivial ideal
(of a certain class of rings related with~$A$) is contained in a prime ideal.
\item The localic Galois group of a Galois extension~$L|k$ is spatial iff
any~$k$-homomorphism~$\sigma : E \to L$ defined on a finite intermediate
extension~$L|E|k$ admits an extension to a~$k$-homomorphism~$L \to L$.
\item The localic real line is spatial iff every open covering of the
metric space~$[0,1]$ of Dedekind reals has a Lebesgue number (that is, if the
metric space~$[0,1]$ is Heine--Borel compact).
\item The statement that the classifying locale of any propositional coherent
theory (that is, a propositional geometric theory in which only finite
disjunctions occur) is spatial is equivalent to the Boolean Prime Ideal
Theorem~\BPIT.
\end{enumerate}

In this picture, the opens of a locale can be regarded as finite approximations
to the ideal points which potentially belong to them. For instance, the
open~$[\sigma_{\sqrt{2},-\sqrt{2}} \wedge \sigma_{\sqrt{3},\sqrt{3}}]$
of the localic Galois group of~$\overline{\QQ}|\QQ$ can be regarded as an
approximation to a certain automorphism of~$\overline{\QQ}$ about
which~$\sqrt{2} \mapsto -\sqrt{2}$ and~$\sqrt{3} \mapsto \sqrt{3}$ is known. Over time, this
approximation might be refined to a smaller open, one which contains more
information about the purported automorphism, and eventually we might actually
obtain an automorphism which is defined on all of~$\overline{\QQ}$. \emph{The opens
make good constructive sense even if the points remain elusive.}

The results mentioned above explain why we can, in classical mathematics, where all of the stated
conditions are satisfied, blithely and fruitfully use the topological
spectrum of a ring, the topological Galois group or the topological space of
reals. In classical mathematics, employing the localic counterparts is merely
for convenience or for aesthetic reasons; in constructive mathematics, the
localic replacements (or similar replacements in other pointfree approaches to
topology) are vital.

\begin{remark}The case of the locale of surjections~$\NN \twoheadrightarrow
\RR$ is slightly different in that there, even the strong existence axioms of
classical mathematics do not suffice to materialize a point.
\end{remark}

@@

Definition~\ref{defn:presheaf}, as well as any other definition in this section,
also makes sense for topological spaces instead of locales, since only the
notion of opens and their inclusion relation is used. The elements of~$F(U)$ are
called \emph{sections of~$F$ over~$U$} and the maps~$(\cdot)|^U_V$
are called \emph{restriction maps}. Elements of~$F(\top)$ are called
\emph{global sections}.

@@

% XXX Introduction...

Let~$X$ be a locale. A \emph{formula over an open~$U \in \O(X)$} is a
first-order formula (made up using~${=}\ {\top}\ {\wedge}\ {\bot}\ {\vee}\
{\Rightarrow}\ {\forall}\ {\exists}$) over the signature which has one sort for
each sheaf~$F$, one constant symbol of sort~$F$ for each section~$s \in F(U)$,
one function symbol~$f : F \to G$ for each morphism of sheaves, and so on. As
usual, we define~$\neg\varphi$ to be an abbreviation of~$(\varphi \Rightarrow
\bot)$.

Table~\ref{table:sheaf-semantics} defines by recursion on the structure of
formulas what it means for a formula~$\varphi$ over an open~$U$ to be
\emph{forced} or \emph{true over~$U$}, written~``$U \models \varphi$''. The
main differences with the perhaps more familiar Kripke semantics are marked.
The sheaf semantics is also called \emph{Kripke--Joyal semantics}.

@@

\begin{definition}\label{defn:sheaf-model}
A \emph{sheaf model} of a geometric theory~$\TT$ over a locale~$X$ consists of
\begin{enumerate}
  \item a sheaf~$\brak{S}$ on~$X$ for each sort~$S$,
  \item a morphism~$\brak{f} : \brak{S_1} \times \cdots \times \brak{S_n} \to
  \brak{T}$ of sheaves for each function symbol~$f : S_1 \times \cdots \times S_n \to T$ and
  \item a subsheaf~$\brak{R} \subseteq \brak{S_1} \times \cdots \times \brak{S_n}$
  for each relation symbol~$R \hookrightarrow S_1 \times \cdots \times S_n$
\end{enumerate}
such that the axioms of~$\TT$ are validated under the sheaf semantics.
\end{definition}

\begin{example}The sheaf~$\C$ of continuous real-valued functions together with
appropriate morphisms is a sheaf model of the theory of rings. Any sheaf can be
regarded as a model of the theory of objects.\end{example}

\begin{example}\label{ex:sets-as-sheaf-models}
Any set-based model of a geometric theory~$\TT$ gives rise to a
sheaf model over the one-point locale, by employing the constant sheaf
construction of XXX. Hence sheaf models are strictly more general than
set-based models.
\end{example}

In the special case of nullary function symbols or relations symbols, by
``$\brak{S_1} \times \cdots \times \brak{S_n}$'' in
Definition~\ref{defn:sheaf-model} we mean the empty product of sheaves, the
\emph{terminal sheaf}~$1$ with~$1(U) = \{\star\}$ for every open~$U$ of~$X$.
Morphisms~$1 \to F$ are in canonical bijection with global sections of~$F$.
Subsheaves~$F \subseteq 1$ are in canonical bijection with the opens of~$X$
(Exercise~\ref{ex:subsheaves-1}).

With this in mind, we can rephrase Definition~\ref{defn:sheaf-model} in the
special case of propositional theories in the following simpler form. A
\emph{sheaf model} of a propositional geometric theory~$\TT$ over a locale~$X$
consists of a family~$(U_\varphi)_\varphi$ of opens of~$X$, one for each
nullary relation symbol~$\varphi$ of~$\TT$, in such a way that the axioms are
satisfied. For instance, that an axiom~``$\alpha \wedge \beta \vdash \bigvee_i
\gamma_i$'' is satisfied means that~$U_\alpha \wedge U_\beta \preceq \bigvee_i
U_{\gamma_i}$.

In other words, a model of a propositional geometric theory~$\TT$ over a
locale~$X$ is the same as a frame homomorphism~$\O(L(\TT)) \to \O(X)$, hence a
locale morphism~$X \to L(\TT)$.

A geometric theory can well have no set-based models, but might still be
consistent and allow for a diverse range of sheaf models on a host of different
locales. In fact, any propositional geometric theory always has at least one,
and very peculiar, sheaf model, namely its \emph{generic model}.

\begin{definition}The \emph{generic model}~$U_\TT$ of a propositional geometric
theory~$\TT$ is the sheaf model on~$L(\TT)$ given by~$U_\varphi \defeq
[\varphi]$ for all nullary relation symbols~$\varphi$.\end{definition}

Under the view of models as locale morphisms, the generic model~$U_\TT$
corresponds to the identity morphism~$L(\TT) \to \L(\TT)$. If a~$\TT$-model~$M$ on
a locale~$X$ is given by a locale morphism~$f : X \to L(\TT)$, we also say
that~``$M$ arises from~$U_\TT$ by pullback along~$f$''.

For instance, there is the \emph{generic Dedekind cut}, the generic model of
the theory~$\TT$ of Dedekind cuts on the classifying locale of~$\TT$, the
localic real line. By Example~\ref{ex:sets-as-sheaves}, any specific Dedekind
real~$x$ (that is, any point~$x$ of~$L(\TT)$) gives rise to a model of~$\TT$ on
the one-point locale~$\pt$, corresponding to the locale morphism~$x : \pt \to
L(\TT)$.

The story of generic models is at the same time tautologous and deep, and these
short paragraphs do not do justice to this intriguing aspect of the interplay
between topology (in the form of locales) and logic (in the form of
propositional geometric theories). We strongly
recommend the leisurely account by Steve Vickers~\ref{XXX} for a fuller
picture.

@@

\begin{exercise}{The localic Galois group of a field extension}%
Let~$L|k$ be a field extension. The \emph{localic Galois group}~$\Gal(L|k)$
is the classifying locale of the theory of automorphisms~$L \to L$ which
fix~$k$, as described on page~\pageref{par:theory-automorphisms}.
\begin{enumerate}
\item Explain how to obtain a morphism~$\Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send an automorphism~$\sigma$ to its inverse~$\sigma^{-1}$.
\item Explain how to obtain a morphism~$\Gal(L|k) \times \Gal(L|k) \to \Gal(L|k)$ which on the
level of points should send a pair~$\langle \sigma, \tau \rangle$ to~$\sigma \circ \tau$.
\item Assume from now on that~$L$ is a \emph{geometric field} in the sense that any
element is zero or invertible. Show that the axiom expressing injectivity can
be dropped without changing the set of provable sequents.
\item Let~$f \in k[X]$ be a monic polynomial such that~$f(X) =
(X-x_1)\cdots(X-x_n)$ over~$L$. Let~$x$ be one of the zeros of~$f$. Show that
the theory shows~$\top \vdash \bigvee_{i=1}^n \sigma_{x,x_i}$.
\item Now furthermore assume that~$L$ is \emph{Galois over}~$k$ in the sense that
any element of~$L$ is the zero of a monic polynomial~$f \in k[X]$ which is
separable (that is,~$(f,f') = (1) \subseteq k[X]$) and splits over~$L$ into
linear factors. Show that the axioms expressing surjectivity can be dropped.
\item Show that the topological space of points of~$\Gal(L|k)$ is canonically
homeomorphic to the topological Galois group of~$L|k$.
\item Show that~$\Gal(L|k)$ is spatial in classical mathematics.
\end{enumerate}
\end{exercise}

@@

\begin{enumerate}
\item Instead of replacing a given ring with another (for instance a ring~$A$ with
one of its stalks~$A_\ppp$ at a minimal prime ideal), we replace a given ring
with a \emph{sheaf of rings}.
\item We then maintain the convenient illusion that we are working with a plain old ring
instead of a sheaf by employing the sheaf semantics.
\end{enumerate}

@@

The locale~$X$ on which this theater plays out is none other than the
\emph{spectrum of~$A$} from algebraic geometry, and the sheaf~$A^\sim$ is the
sheaf of regular functions, also denoted~``$\O_{\Spec(A)}$''. We will give details on its construction in
Section~\ref{sect:origins}, but first want to reflect on the displayed
statements.

@@

Not only does the ring~$A$ have a mirror image as a sheaf, also
any~$A$-module~$M$ does. Its mirror image is the \emph{quasicoherent
sheaf}~$M^\sim$ from algebraic geometry. As with~$A$ and~$A^\sim$, there is a
close connection between~$M$ and~$M^\sim$; for instance, the module~$M$ is
finitely generated over~$A$ iff~$M^\sim$ is finitely generated over~$A^\sim$
(Exercise~\ref{ex:bridging-modules}).

\paragraph{The field property.}
The fact that~$A^\sim$ is a field is where~$A^\sim$ derives its main usefulness
from. While statements~(\ref{item:loc-stable}), (\ref{item:stalks}) and
(\ref{item:local}) are immediate consequences of the way~$A^\sim$ is
constructed (Proposition~\ref{prop:immediate-consequences}), the field property~(\ref{item:field})
requires an actual computation for its verification and was a surprising
discovery by Christopher Mulvey in the 1970s.

Myles Tierney commented around that time that the field property ``is surely important,
though its precise significance is still somewhat obscure---as is the case with
many such nongeometric formulas''~\cite[p.~209]{tierney:spectrum}. Even though
the property has been known for a long time, only recently was it put in a
wider context (recognizing it as a small shadow of algebraic geometry's
\emph{quasicoherence}) and its usefulness for constructive algebra
appreciated~\cite[Sections~3.3 and~11.4]{blechschmidt:phd}.

@@

In exactly the same manner, we can construct the free group over a monoid, the
free~$\ZZ/(2)$-algebra over a ring, the free ring over a set or the free
Heyting algebra over a partially ordered set. This is all well-understood; a
particularly perspicious account in the general setting of partial Horn
theories is due to Erik Palmgren and Steve Vickers~\cite{palmgren-vickers:partial-horn}.

@@

\begin{exercise}{Bridging modules and their induced sheaves}%
\label{ex:bridging-modules}%
Let~$M$ be a module over a ring~$A$.
\begin{enumerate}
\item Show that~$M$ is finitely generated iff~$M^\sim$ is.
\item Show that~$M$ is finite locally free (that is, that there is a
partition~$1 = f_1 + \cdots + f_n$ such that each module~$M[f_i^{-1}]$ is
finite free over~$A[f_i^{-1}]$) iff~$M^\sim$ is finite free.
\item Show that~$M$ is finitely presented iff~$M^\sim$ is.
\end{enumerate}
\end{exercise}

@@

\begin{exercise}{Spatiality of the spectrum}%
Let~$A$ be a ring. Let~$\Spec(A)$ be its spectrum.
\begin{enumerate}
\item Verify that~$\Spec(A)$ is spatial if and only if for any element~$x \in
A$ and any ideal~$\aaa \subseteq A$,
\[ x \in \sqrt{\aaa} \quad\textnormal{if}\quad
\text{for all prime filters~$\fff \subseteq A$ with~$x \in \fff$, $\fff \between \aaa$}, \]
where the symbol~``$\between$'' denotes that the two sets have an element in common.
\item Assuming the law of excluded middle, show that~$\Spec(A)$ is spatial if
and only if for any ideal~$\aaa \subseteq A$ the familiar identity
\[ \sqrt{\aaa} = \bigcap_{\ppp \supseteq \aaa} \ppp \]
holds.
\item Conclude that the spectrum of any ring is spatial iff~\BPIT holds.
\end{enumerate}
\end{exercise}

@@

FUTURE:
* Why the name "geometric"?
